\thetitle

### 2020-11-21 writing/discussion.tex
revision of results

\subsection{Why does the data look like this?}

The figures in the last section raise a key question: why does blame fail in our dataset?
We analyze the reasons for failure separately for each mode of the rational programmer.

Breaking down the reasons for failure for all scenarios in which Natural is unsuccessful reveals two causes.
First, most of the failing scenarios begin with a runtime exception, which the rational programmer can use to type components until blame becomes available.
In these failing cases (there are 1708), however, blame never becomes available; the exception information points to the wrong component and the trail ends in failure.
A second, much smaller set (40 scenarios) do produce a runtime type error blaming some component, but the blame identifies a non-buggy typed component.
After careful inspection of these scenarios, we are confident that they reveal a bug in the implementation of Typed Racket.

Transient's blame failures break down in a similar way.
Like Natural, most failures occur with unhelpful exception information without ever getting blame (1851 scenarios).
However, Transient also has a substantial number of failures that occur because the programs hit the resource limits (timed out or ran out of memory) and had to be killed (some 770 scenarios).
Additionally, there are nearly a thousand cases where Transient's checks catch a problem, but the blame set reported is empty.
We discuss each of these last two failure categories in detail in~\ref{sec:threat:transient}.

\subsection{High level conclusions}

The analysis in the last section indicates a number of interesting conclusions about blame in the gradual typing setting.
First, figure~\ref{fig:avo-matrix} makes it clear that any kind of dynamic checking has a large positive impact over Erasure, and this is true regardless of whether the rational programmer uses blame or stacktraces.
Indeed, the bottom row of figure~\ref{fig:avo-matrix} illustrates that Erasure almost never offers more useful information than any other mode.
Nonetheless, figure~\ref{fig:effort-table} does show that even if erasure often fails to provide helpful information, it provides the rational programmer with a quick answer.

Figure~\ref{fig:avo-matrix} also indicates that blame helps significantly over just using exception information.
Both Natural and Transient's blame modes improve over their corresponding exception modes by about ten percent.
This illustrates that blame tracking offers value for locating the source of mistakes beyond the value provided by the dynamic checks alone.
At only ten percent, however, the amount of added value suggests that foregoing blame tracking in order to obtain performance gains may not hurt debugging all too much.

Finally, the results show that while Natural clearly offers the most utility compared to the other modes, Transient is not too far behind.
Natural is more useful than every other mode in figure~\ref{fig:avo-matrix}, and by a larger proportion than any other mode.
That said, the same can be said of Transient's two blame modes as compared to every other mode but Natural.
Furthermore, the difference in usefulness between Natural and Transient's blame modes is quite small across the board.
As we point out in~\ref{sec:mutate}, however, we should note that our selection of scenarios for analysis introduces bias toward Erasure and Transient.

\subsection{Threats to validity}

@@

\subsection{Threat: Is our Transient Correct?}
\label{sec:threat:transient}
%% purpose: talk about transient-blame implementation .... what did we do,
%%  how did we test, what threats remain?
%%
%% the benchmark timeouts look like a threat, but this section shows that
%%  we've looked carefully at those

>>>>

The analysis in the previous section suggests a number of interesting
high-level conclusions about blame in the gradual typing setting.  First,
run time type checks have a large positive impact. This is the case
regardless of whether these checks issue blame or throw plain exceptions.
Second, in many cases type checks that issue blame are more helpful than
those that do not. However, our results indicate that in other cases blame
is not critical and thus it is worth further investigation whether
disabling blame tracking for the sake of performance may be   an
acceptable trade off. Third, from the three different blame strategies the
Natural approach of Typed Racket fares better than the Transient approach
of Reticulated Python but by a small margin. Since Natural offers complete
and sound path-based blame while Transient incomplete but sound heap-based
blame~\cite{gfd-oopsla-2019}, our results call for a deeper understanding
of the relative strengths of the two models for blame.  Fourth, given that
Transient's sound but shallow run time type checks do not seem to hamper
debugging, a version of Typed Racket that disables some of its wrappers
may offer an answer to the performance issues of the Natural semantics. 
Fifth, just as for Typed Racket, the experiment points to possible
improvements for Reticulated Python. In particular, the experimental 
data indicates that the usefulness of transient blame does not depend on
whether  the rational programmer follows first or the last element from a
Transient blame set. Hence, given the performance issues of Transient we discuss further on, 
a possible optimization is to limit the size of blame sets
using timestamps.

There are a number of threats to the

\subsection{Threat: Is the Rational Programmer Realistic?}

@@

\subsection{Threat: Is our Transient Correct?}
\label{sec:threat:transient}

### 2020-11-21 writing/results.tex
Lots of minor changes and rewordings
revision of results

\section{What We Discovered}

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} using a machine with two Intel Xeon Gold 6258R processors and 500GB of memory.
Each debugging scenario had a timeout limit of 4 minutes and a
memory limit of 6GB. The complete experiment required
approximately 270 hours of compute hours on our machine.

@@

A third notable piece of information from
the figure is that every mode has failed debugging scenarios, not just
Erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program, or only typed modules. This situation
can even happen immediately, which explains failures with 0 effort. 
While the most
failures to debug scenarios follow this pattern, a few do not. The next section takes a
closer look at these unexpected failures of the rational programmer as part of the lessons we
extract from our experimental results.

>>>>

\section{What We Discovered}
\label{sec:results}

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} using a machine with two Intel Xeon Gold 6258R processors and 500GB of memory.
Each debugging scenario had a timeout limit of 4 minutes and a
memory limit of 6GB. The complete experiment required
over 30,000 hours of compute hours on our machine.

@@

A third notable piece of information from
the figure is that every mode has failed debugging scenarios, not just
Erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program, or only typed modules. This situation
can even happen immediately, which explains failures with 0 effort.

While most failures to debug scenarios follow the above pattern, a few do not.
Breaking down the reasons for failure for Natural blame (1748 in total)
reveals an additional cause. For a small set of
debugging scenarios (40), Natural produces a run time type error
blaming a non-buggy already typed
component. We tracked down all these cases to known open issues with Typed
Racket and class contracts.

In Transient, similar to Natural,
most failures are due to unhelpful exception information (1851 for both
Transient first and last blame).  
However, Transient also has a substantial
number of failures because scenarios hit the time and memory
limits of our experiment (~ 770 scenarios).  Additionally, there are nearly a 1000 cases where
Transient reports an empty blame set which leaves the rational programmer
without hints about how to proceed.
In~\ref{sec:threat:transient}, we discuss how both of these causes of
failure for transient affect our experiment.
